Forecasting the 2024 Election: Trump Dominates the South, Harris Leads in the Midwest

### 2024-10-20 paper/paper.qmd
committing formatting changes to paper.qmd
committing changes to paper.qmd

The results of our model reveal clear geographic divides in voter sentiment, with Donald Trump predicted to win in many southern and midwestern states, while Kamala Harris is expected to perform well in the Midwest and on the East Coast. The predictions align with historical voting patterns, with Trump’s strongest support coming from traditionally conservative regions and Harris garnering more support in Democratic-leaning areas. Notably, several battleground states, such as Nevada and North Carolina, show a close contest, reflecting the competitive nature of the 2024 election. These results suggest that the election may hinge on the outcomes in these key states, where even small shifts in polling percentages could influence the final result. The predicted probabilities in states like Michigan, Virginia, and Pennsylvania indicate a strong chance for Harris, while Trump's significant lead in states like Florida and Missouri solidifies his dominance in certain regions. Overall, the model provides valuable insights into the competitive landscape, highlighting the states that could be decisive in determining the election outcome.

@@

The logistic regression model used in this analysis successfully captures the relationship between polling percentages and the likelihood of a candidate winning a state. The coefficients for the predictor variables (polling percentages for Trump and Harris) indicate a strong relationship between candidate support and the probability of winning. The model outputs probabilities for each state, which are useful for gauging the certainty of predictions, especially in closely contested states. The model’s performance is reinforced by the robustness of the results in states where one candidate has a clear polling lead, providing accurate predictions. However, states with narrower margins reflect greater uncertainty, which is critical in understanding election dynamics. Convergence diagnostics and model checking were conducted to ensure reliable estimates, with further details and alternative models provided in [Appendix -@sec-model-details].

@@

Despite the model's predictive power, there are several weaknesses that should be noted. First, polling data itself is prone to biases, such as underrepresentation of certain demographic groups or inaccuracies in sampling. These issues can distort the predictions, particularly in states with fewer or lower-quality polls. Additionally, voter turnout and last-minute shifts in public opinion are not accounted for, which can significantly alter the outcome, especially in close races. The model also assumes that polling percentages directly translate to votes, overlooking other factors like voter mobilization efforts and campaign dynamics that may influence election outcomes.

To improve the accuracy and reliability of election forecasts, future iterations of the model could incorporate additional variables beyond polling data. Factors such as voter turnout models, economic conditions, and social media sentiment analysis may provide a more holistic view of voter behavior. Additionally, real-time polling data integration could allow for dynamic updates to predictions as the election nears, improving the timeliness and accuracy of forecasts. Moreover, regional adjustments to account for differences in polling methodology or demographic shifts would help to address some of the biases inherent in polling data. Expanding the model to include these aspects would offer a more comprehensive understanding of the election landscape and enhance the precision of predictions.

>>>>

Placeholder - The results of our model reveal clear geographic divides in voter sentiment, with Donald Trump predicted to win in many southern and midwestern states, while Kamala Harris is expected to perform well in the Midwest and on the East Coast. The predictions align with historical voting patterns, with Trump’s strongest support coming from traditionally conservative regions and Harris garnering more support in Democratic-leaning areas. Notably, several battleground states, such as Nevada and North Carolina, show a close contest, reflecting the competitive nature of the 2024 election. These results suggest that the election may hinge on the outcomes in these key states, where even small shifts in polling percentages could influence the final result. The predicted probabilities in states like Michigan, Virginia, and Pennsylvania indicate a strong chance for Harris, while Trump's significant lead in states like Florida and Missouri solidifies his dominance in certain regions. Overall, the model provides valuable insights into the competitive landscape, highlighting the states that could be decisive in determining the election outcome.

@@

Placeholder - The logistic regression model used in this analysis successfully captures the relationship between polling percentages and the likelihood of a candidate winning a state. The coefficients for the predictor variables (polling percentages for Trump and Harris) indicate a strong relationship between candidate support and the probability of winning. The model outputs probabilities for each state, which are useful for gauging the certainty of predictions, especially in closely contested states. The model’s performance is reinforced by the robustness of the results in states where one candidate has a clear polling lead, providing accurate predictions. However, states with narrower margins reflect greater uncertainty, which is critical in understanding election dynamics. Convergence diagnostics and model checking were conducted to ensure reliable estimates, with further details and alternative models provided in [Appendix -@sec-model-details].

@@

Placeholder - Despite the model's predictive power, there are several weaknesses that should be noted. First, polling data itself is prone to biases, such as underrepresentation of certain demographic groups or inaccuracies in sampling. These issues can distort the predictions, particularly in states with fewer or lower-quality polls. Additionally, voter turnout and last-minute shifts in public opinion are not accounted for, which can significantly alter the outcome, especially in close races. The model also assumes that polling percentages directly translate to votes, overlooking other factors like voter mobilization efforts and campaign dynamics that may influence election outcomes.

Placeholder - To improve the accuracy and reliability of election forecasts, future iterations of the model could incorporate additional variables beyond polling data. Factors such as voter turnout models, economic conditions, and social media sentiment analysis may provide a more holistic view of voter behavior. Additionally, real-time polling data integration could allow for dynamic updates to predictions as the election nears, improving the timeliness and accuracy of forecasts. Moreover, regional adjustments to account for differences in polling methodology or demographic shifts would help to address some of the biases inherent in polling data. Expanding the model to include these aspects would offer a more comprehensive understanding of the election landscape and enhance the precision of predictions.
